% Problem 2-2

\begin{itemize}
  \item[\textbf{a.}] We need to prove a loop invariant for each loop
    in the algorithm.
  \item[\textbf{b.}]
    \begin{proof}
      At the start of each iteration of the \textbf{for} loop of lines 2-4,
      $A'[1\ltwodots i]$ consists of elements sorted in ascending order.
      \begin{description}
        \item[Initialisation:] For the first loop of lines 2-4, $A'[1
          \ltwodots i]$
          is a one-element array, $A'[1]$, which is (trivially) sorted.
        \item[Maintenance:] After each \textbf{for} loop, the first out-of-order
          element from $A.\mathit{length}$ down to $i + 1$ has been swapped with
          each element before it until it is in the correct position in
          $A'[1 \ltwodots i]$.
        \item[Termination:] The \textbf{for} loop terminates when $j == i + 1$,
          so that, at that point, $A'[i] < A'[i + 1]$ and the loop invariant
          holds for $A'[1 \ltwodots i + 1]$.
      \end{description}
    \end{proof}
  \item[\textbf{c.}]
    \begin{proof}
      For each iteration of the \textbf{for} loop of lines 1-4, $A'[1
      \ltwodots i + 1]$
      is sorted in ascending order.
      \begin{description}
        \item[Initialisation:] At initialisation, $i = 1$, so $A'[1
          \ltwodots i + 1]$
          is $A'[1 \ltwodots 2]$ and the termination of \textbf{b.} guarantees
          that this will be sorted.
        \item[Maintenance:] Each execution of the \textbf{for} loop of lines 2-4
          sorts $A'[1 \ltwodots i + 1]$, so the loop invariant is maintained.
        \item[Termination:] The loop terminates when $i ==
          A.\mathit{length} - 1$,
          so $A'[1 \ltwodots i + 1] = A'[1 \ltwodots n]$ has been sorted.
      \end{description}
    \end{proof}
  \item[\textbf{d.}] The worst-case running time occurs when $A$ is
    sorted in descending
    order, so the \textbf{for} loop of lines 1-4 executes $n - 1$
    times and the \textbf{for}
    loop of lines 2-4 executes $n - i + 1$ times. Thus we have
    $(n - 1)(n - i + 1) = \Theta(n^2)$ worst-case running time,
    equivalent to that of
    \textsc{Insertion-Sort}.
\end{itemize}

The implementation can be seen in the following file:

\path{rs/clrs_algorithms/src/sorting.rs}.
